---
title: "Exp. 2: pre-registered results"
shorttitle: "Exp. 2: pre-registered results"

author: 
  - name: "Maya Schipper"
    affiliation: "1"
  - name: "Matan Mazor"
    affiliation: "1"

affiliation:
  - id: "1"
    institution: "University of Oxford"

floatsintext: no
linenumbers: yes
draft: no
mask: no

figurelist: no
tablelist: no
footnotelist: no

classoption: "man"
output: papaja::apa6_word
bibliography: "references.bib"
---

```{r global_options, include=FALSE, cache=F}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, cache=T, sanitize = T)
```

```{r setup, include = FALSE, cache=F}

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'osfr',
    'tidyverse',
    'psych'
  ), "2024-12-01"
)

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)

```

```{r download_load_and_format_data, echo=FALSE, cache=TRUE}

source("../analysis/downloadFromOSF.R")
source("../analysis/loadAndPreprocessData.R")
```

### Participants

`r prospective.raw$subj_id%>%unique()%>%length()` participants took part in the letter-detection task. Participants were excluded if their accuracy fell below 50% (0 participants). We also excluded participants if they had extremely fast (`r prospective.too_fast%>%length()` participants) or slow (`r prospective.too_fast%>%length()` participants) reaction times in this task (below 100 milliseconds or above 7 seconds in more than 25% of the trials). Trials with reaction times below 100 milliseconds (`r (prospective.raw%>%filter(RT<100)%>%nrow()/prospective.raw%>%filter(RT>0)%>%nrow())%>%printnum()` of all trials) or above 7 seconds (`r (prospective.raw%>%filter(RT>7000)%>%nrow()/prospective.raw%>%filter(RT>0)%>%nrow())%>%printnum()` of all trials) were excluded from the reaction time analysis.

### Hypotheses

This study is designed to test whether expected visibility impacts participants’ decisions about presence and absence.

```{r H1, echo=FALSE, cache=TRUE}
response_RT <- prospective.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & correct == TRUE) %>%  
    dplyr::group_by(subj_id, resp) %>%
    dplyr::summarise(RT = median(RT)) %>%
    tidyr::pivot_wider(names_from = resp, values_from = RT) %>%
    dplyr::mutate(response_RT = `TRUE` - `FALSE`)
```

*Hypothesis 1 (PRESENCE/ABSENCE RESPONSE TIME)*: We tested the null hypothesis that response times are similar for target-present and target-absent responses, aiming to replicate the finding that decisions about the target-absence are slower than decisions about target-presence [@mazor2021].

We performed paired t-test on the following individual-level contrast from correct trials: $median({RT}_{P})-median({RT}_{A})$.

This revealed a significant difference in response times for target-present and target-absent responses (`r apa_print(response_RT%>%pull(response_RT)%>%t.test())$full_result`), with overall slower response times for decisions about absence.

```{r H2, echo=FALSE, cache=TRUE}
RT_size_present <- prospective.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & present == 1 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(RT = median(RT))%>%
    tidyr::pivot_wider(names_from = size, values_from = RT) %>%
    dplyr::mutate(RT_size_present = big - small)
```

*Hypothesis 2 (SIZE EFFECT ON REACTION TIME IN PRESENCE)*: We tested the null hypothesis that target-present response times are similar when the display is big and small.

We performed a paired t-test on the following individual-level contrast from correct trials: $median({RT}_{P, big})-median({RT}_{P, small})$ where P stands for present.

This revealed a significant difference in response times between big and small stimuli in target-present decisions (`r apa_print(RT_size_present%>%pull(RT_size_present)%>%t.test())$full_result`), with slower response times for big stimuli.

```{r H3, echo=FALSE, cache=TRUE}
RT_size_absent <- prospective.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & present == 0 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(RT = median(RT))%>%
    tidyr::pivot_wider(names_from = size, values_from = RT) %>%
    dplyr::mutate(RT_size_absent = big - small)
```

*Hypothesis 3 (SIZE EFFECT ON REACTION TIME IN ABSENCE)*: We tested the null hypothesis that target-absent response times are similar when the display is big and small.

We performed a paired t-test on the following individual-level contrast from correct trials: $median({RT}_{A, big})-median({RT}_{A, small})$ where A stands for absent.

This revealed a null effect of stimulus size on response time in target-absent decisions (`r apa_print(RT_size_absent%>%pull(RT_size_absent)%>%t.test())$full_result`), with a small but insignificant increase in response times for small stimuli.

```{r H4, echo=FALSE, cache=TRUE}
RT_size_resp <- inner_join(
  RT_size_present,
  RT_size_absent,
  by = "subj_id") %>%
  mutate(RT_size_resp = RT_size_present - RT_size_absent)
```

*Hypothesis 4 (SIZE RESPONSE INTERACTION ON REACTION TIME)*: We tested the null hypothesis that the effect of stimulus size on reaction time is similar in target-absent and target-present responses.

We performed a group-level t-test on a subject-level contrast in correct trials only: $(median({RT}_{P,small})-median({RT}_{P,big}))-(median({RT}_{A,small}) - median({RT}_{A,big}))$ where P and A stand for present and absent respectively.

This revealed a non-significant interaction between stimulus size and target presence (`r apa_print(RT_size_resp%>%pull(RT_size_resp)%>%t.test())$full_result`).

```{r H5, echo=FALSE, cache=TRUE}
accuracy_size <- prospective.df %>%
  dplyr::group_by(subj_id, size) %>%
  dplyr::summarise(
    hit_rate =(sum(correct & present)+0.5)/(sum(present)+1),
    fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
    dprime = qnorm(hit_rate)-qnorm(fa_rate),
    criterion = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))

dprime <- accuracy_size %>%
  dplyr::select(subj_id,size,dprime) %>%
  tidyr::pivot_wider(names_from=size, values_from=dprime) %>%
  dplyr::mutate(dprime_size=big-small)
```

*Hypothesis 5 (SENSITIVITY)*: We tested the null hypothesis that perceptual sensitivity (measured as $d\prime=z(H)-z(F)$) is equal as a function of the size of the stimulus. To allow the extraction of d’ for participants who committed no false-alarms or misses, we added 0.5 to miss, hit, false-alarm and correct rejection counts [@snodgrass2002].

We performed a paired t-test on the following subject-level contrast: $d\prime_{big} - d\prime_{small}$.

We found a significant effect of stimulus size on $d'$ (`r apa_print(dprime%>%pull(dprime_size)%>%t.test())$full_result`), with increased sensitivity for small stimuli.

```{r H6, echo=FALSE, cache=TRUE}
criterion <- accuracy_size %>%
  dplyr::select(subj_id,size,criterion) %>%
  tidyr::pivot_wider(names_from=size, values_from=criterion) %>%
  dplyr::mutate(criterion_size=big-small)
```

*Hypothesis 6 (CRITERION)*: We tested the null hypothesis that decision criterion (measured as $c=-0.5(z(H)+z(F))$) is unaffected by the size of the display. To allow the extraction of a decision criterion for participants who committed no false-alarms or misses, we added 0.5 to miss, hit, false-alarm and correct rejection counts [@snodgrass2002].

We performed a paired t-test on the following subject-level contrast: $c_{big} - c_{small}$.

We found a significantly more conservative criterion for big stimuli (`r apa_print(criterion%>%pull(criterion_size)%>%t.test())$full_result`).

```{r H7, echo=FALSE, cache=TRUE}
response_confidence <- prospective.df %>%
    dplyr::filter(correct == TRUE & confidence_time == "after") %>%
    dplyr::group_by(subj_id, resp) %>%
    dplyr::summarise(confidence=mean(confidence)) %>%
    tidyr::pivot_wider(names_from = resp, values_from = confidence) %>%
    dplyr::mutate(response_confidence = `TRUE` - `FALSE`)
```

*Hypothesis 7 (PRESENCE/ABSENCE CONFIDENCE RATINGS)*: We tested the null hypothesis that post-decisional confidence ratings are similar for target-present and target-absent responses, aiming to replicate the finding that subjective confidence levels are lower for decisions about absence than decisions about presence [@mazor2021].

We performed a paired t-test on the the following subject-level contrast from correct trials: $mean(confidence_{present}) - mean(confidence_{absent})$

This revealed a significant difference in confidence ratings for decisions about presence and absence (`r apa_print(response_confidence%>%pull(response_confidence)%>%t.test())$full_result`), with higher confidence ratings for present responses.

```{r H8, echo=FALSE, cache=TRUE}
conf_size_present <- prospective.df %>%
    dplyr::filter(present == 1 & correct == TRUE & confidence_time == "after") %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(confidence = mean(confidence)) %>%
    tidyr::pivot_wider(names_from = size, values_from = confidence) %>%
    dplyr::mutate(conf_size_present = big-small)
```

*Hypothesis 8 (SIZE EFFECT ON CONFIDENCE IN PRESENCE)*: We tested the null hypothesis that post-decisional target-present confidence ratings are similar when the stimulus is small or big.

We performed a paired t-test on the following individual-level contrast from correct trials: $mean(confidence_{P, big}) - mean(confidence)_{P, small}$ where P stands for present.

This revealed no difference in confidence ratings for big and small stimuli in target-present responses (`r apa_print(conf_size_present%>%pull(conf_size_present)%>%t.test())$full_result`), with higher confidence ratings for small stimuli.

```{r H9, echo=FALSE, cache=TRUE}
conf_size_absent <- prospective.df %>%
    dplyr::filter(present == 0 & correct == TRUE & confidence_time == "after") %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(confidence = mean(confidence)) %>%
    tidyr::pivot_wider(names_from = size, values_from = confidence) %>%
    dplyr::mutate(conf_size_absent = big-small)
```

*Hypothesis 9 (SIZE EFFECT ON CONFIDENCE IN ABSENCE)*: We tested the null hypothesis that post-decisional target-absent confidence ratings are similar when the stimulus is small or big.

We performed a paired t-test on the following individual-level contrast from correct trials: $mean(confidence_{A, big}) - mean(confidence)_{A, small}$ where A stands for absent.

This revealed a significant effect of stimulus size on confidence ratings in target-absent responses (`r apa_print(conf_size_absent%>%pull(conf_size_absent)%>%t.test())$full_result`), with higher confidence ratings for big stimuli.

```{r H10, echo=FALSE, cache=TRUE}
conf_size_resp <- inner_join(
  conf_size_present,
  conf_size_absent,
  by = "subj_id") %>%
  dplyr::mutate(conf_size_resp = conf_size_present - conf_size_absent)
```

*Hypothesis 10 (SIZE RESPONSE INTERACTION ON CONFIDENCE)*: We tested the null hypothesis that the effect of stimulus size on confidence ratings is similar in target-present and target-absent responses.

We performed a group-level t-test on a subject-level contrast in correct trials only: $(mean({confidence}_{P,small})-mean({confidence}_{P,big}))-(mean({confidence}_{A,small}) -mean({confidence}_{A,big}))$ where P and A stand for present and absent respectively.

This revealed a significant interaction effect between stimulus size and target presence on subjective confidence levels (`r apa_print(conf_size_resp%>%pull(conf_size_resp)%>%t.test())$full_result`).

```{r H11, echo=FALSE, cache=TRUE}
pros_conf <- prospective.df %>%
  dplyr::filter(correct==TRUE & confidence_time == "before") %>%
  dplyr::group_by(subj_id, size) %>%
  dplyr::summarise(confidence = mean(confidence)) %>%
  tidyr::pivot_wider(names_from = size, values_from = confidence) %>%
  dplyr::mutate(pros_conf = `big` - `small`)
t.test(pros_conf$pros_conf)
```

*Hypothesis 11 (SIZE EFFECT ON PROSPECTIVE CONFIDENCE)*: We tested the null hypothesis that prospective confidence ratings are similar when the stimulus is expected to be small or big.
This was tested using a paired t-test on the mean individual-level confidence ratings from correct trials only.

We performed a paired t-test on the following individual-level contrast from correct trials: $mean(confidence_{prospective, big}) - mean(confidence)_{prospective, small}$.

This revealed a significant effect of stimulus size on prospective confidence ratings (`r apa_print(pros_conf%>%pull(pros_conf)%>%t.test())$full_result`), with higher confidence ratings for big stimuli.

```{r H12, echo=FALSE, cache=TRUE}
diff_wide <- pros_conf %>%
  dplyr::select(subj_id,pros_conf) %>%
  merge(conf_size_present %>%
          dplyr::select(subj_id,conf_size_present)) %>%
  merge(conf_size_absent %>%
          dplyr::select(conf_size_absent)) %>%
  merge(RT_size_present %>%
          dplyr::select(subj_id,RT_size_present)) %>%
  merge(RT_size_absent %>%
          dplyr::select(subj_id, RT_size_absent)) 

cor_pros_absent <- cor.test(diff_wide$pros_conf, diff_wide$conf_size_absent, method = "pearson")

```

*Hypothesis 12 (SIZE EFFECT CORRELATION: PROSPECTIVE AND ABSENCE)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is uncorrelated with the effect of stimulus size on post-decisional confidence in absence. The two effects were uncorrelated across participants (`r apa_print(cor_pros_absent)$full_result`)

```{r H13, echo=FALSE, cache=TRUE}
cor_pros_present <- cor.test(diff_wide$pros_conf, diff_wide$conf_size_present, method = "pearson")
```

*Hypothesis 13 (SIZE EFFECT CORRELATION: PROSPECTIVE AND PRESENCE)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is uncorrelated with the effect of stimulus size on post-decisional confidence in presence. The two effects were uncorrelated across participants (`r apa_print(cor_pros_present)$full_result`)


```{r H14, echo=FALSE, cache=TRUE}
cor_pres_abs <- cor.test(diff_wide$conf_size_present, diff_wide$conf_size_absent, method = "pearson")
included_N <- diff_wide$subj_id%>%unique()%>%length()
steiger = r.test(included_N,cor_pros_absent$estimate,cor_pros_present$estimate,cor_pres_abs$estimate)

```

*Hypothesis 14 (DIFFERENCE IN CORRELATIONS)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is similarly correlated with the effect of stimulus size on post-decisional confidence in presence and in absence.
This was tested with a Steiger’s t-test for dependent correlations [@@steiger_tests_1980], giving rise to a non-significant ($t=$`r printnum(steiger$t)`, $p$ `r print_p(steiger$p, add_equals=T)`).

```{r H15, echo=FALSE, cache=TRUE}
cor_pros_rtabsent <- cor.test(diff_wide$pros_conf, diff_wide$RT_size_absent, method = "pearson")
```

*Hypothesis 15 (SIZE EFFECT CORRELATION: PROSPECTIVE CONFIDENCE AND ABSENCE RT)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is uncorrelated with the effect of stimulus size on target-absent response times in post-decisional confidence trials. The two were uncorrelated (`r apa_print(cor_pros_rtabsent)$full_result`)

```{r H16, echo=FALSE, cache=TRUE}
cor_pros_rtpresent <- cor.test(diff_wide$pros_conf, diff_wide$RT_size_present, method = "pearson")
```

*Hypothesis 15 (SIZE EFFECT CORRELATION: PROSPECTIVE CONFIDENCE AND PRESENCE RT)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is uncorrelated with the effect of stimulus size on target-present response times in post-decisional confidence trials. The two were uncorrelated (`r apa_print(cor_pros_rtpresent)$full_result`)

```{r H17, echo=FALSE, cache=TRUE}
cor_rtpres_rtabs <- cor.test(diff_wide$RT_size_present, diff_wide$RT_size_absent, method = "pearson")
steiger_RT <- r.test(included_N,cor_pros_rtabsent$estimate,cor_pros_rtpresent$estimate,cor_rtpres_rtabs$estimate) 

```

*Hypothesis 17 (DIFFERENCE IN CORRELATIONS)*: We tested the null hypothesis that the effect of stimulus size on prospective confidence is similarly correlated with the effect of stimulus size on RT in presence and in absence.
This was tested with a Steiger’s t-test for dependent correlations [@@steiger_tests_1980], giving rise to a non-significant ($t=$`r printnum(steiger_RT$t)`, $p$ `r print_p(steiger_RT$p, add_equals=T)`).
\newpage

## References
