---
title: "Expected visibility"
shorttitle: "Expected visibility"

author: 
  - name: "Maya Schipper"
    affiliation: "1"
  - name: "Matan Mazor"
    affiliation: "1"

affiliation:
  - id: "1"
    institution: "University of Oxford"

floatsintext: no
linenumbers: yes
draft: no
mask: no

figurelist: no
tablelist: no
footnotelist: no

classoption: "man"
output: papaja::apa6_word
bibliography: "references.bib"
---

```{r global_options, include=FALSE, cache=F}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, cache=T, sanitize = T)
```

```{r setup, include = FALSE, cache=F}

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse',
    'BayesFactor',# for Bayesian t test
    'ggrepel', # for geom_label_repel
    'lmerTest',
    'osfr'
  ), "2024-12-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)

```

```{r download_load_and_format_data, echo=FALSE, cache=TRUE}

source("../analysis/downloadFromOSF.R")
source("../analysis/loadAndPreprocessData.R")
size.df <- size.df %>% filter(
  test_part %in% c('test1','test2')
)

```

### Participants

`r size.raw_df$subj_id%>%unique()%>%length()` participants took part in the letter-detection task. Participants were excluded if their accuracy fell below 50% (0 participants). We also excluded participants if they had extremely fast (`r size.too_fast%>%length()` participants) or slow (`r size.too_fast%>%length()` participants) reaction times in this task (below 100 milliseconds or above 7 seconds in more than 25% of the trials). Trials with reaction times below 100 milliseconds (`r (size.raw_df%>%filter(RT<100)%>%nrow()/size.raw_df%>%filter(RT>0)%>%nrow())%>%printnum()` of all trials) or above 7 seconds (`r (size.raw_df%>%filter(RT>7000)%>%nrow()/size.raw_df%>%filter(RT>0)%>%nrow())%>%printnum()` of all trials) were excluded from the reaction time analysis.

### Hypotheses

This study is designed to test whether expected visibility impacts participants’ decisions about presence and absence.

```{r H1, echo=FALSE, cache=TRUE}
response_RT <- size.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & correct == TRUE) %>%  
    dplyr::group_by(subj_id, resp) %>%
    dplyr::summarise(RT = median(RT)) %>%
    tidyr::pivot_wider(names_from = resp, values_from = RT) %>%
    dplyr::mutate(response_RT = `TRUE` - `FALSE`)
```

*Hypothesis 1 (PRESENCE/ABSENCE RESPONSE TIME)*: We tested the null hypothesis that response times are similar for target-present and target-absent responses, aiming to replicate the finding that decisions about the target-absence are slower than decisions about target-presence [@mazor2021].

We performed paired t-test on the following individual-level contrast from correct trials: $median({RT}_{P})-median({RT}_{A})$.

This revealed a significant difference in response times for target-present and target-absent responses (`r apa_print(response_RT%>%pull(response_RT)%>%t.test())$full_result`), with overall slower response times for decisions about absence.

```{r H2, echo=FALSE, cache=TRUE}
RT_size_present <- size.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & present == 1 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(RT = median(RT))%>%
    tidyr::pivot_wider(names_from = size, values_from = RT) %>%
    dplyr::mutate(RT_size_present = big - small)
```

*Hypothesis 2 (SIZE EFFECT ON REACTION TIME IN PRESENCE)*: We tested the null hypothesis that target-present response times are similar when the display is big and small.

We performed a paired t-test on the following individual-level contrast from correct trials: $median({RT}_{P, big})-median({RT}_{P, small})$ where P stands for present.

This revealed a significant difference in response times between big and small stimuli in target-present decisions (`r apa_print(RT_size_present%>%pull(RT_size_present)%>%t.test())$full_result`), with slower response times for big stimuli.

```{r H3, echo=FALSE, cache=TRUE}
RT_size_absent <- size.df %>%
    dplyr::filter(RT > 100 & RT < 7000 & present == 0 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(RT = median(RT))%>%
    tidyr::pivot_wider(names_from = size, values_from = RT) %>%
    dplyr::mutate(RT_size_absent = big - small)
```

*Hypothesis 3 (SIZE EFFECT ON REACTION TIME IN ABSENCE)*: We tested the null hypothesis that target-absent response times are similar when the display is big and small.

We performed a paired t-test on the following individual-level contrast from correct trials: $median({RT}_{A, big})-median({RT}_{A, small})$ where A stands for absent.

This revealed a null effect of stimulus size on response time in target-absent decisions (`r apa_print(RT_size_absent%>%pull(RT_size_absent)%>%t.test())$full_result`), with a small but insignificant increase in response times for small stimuli.

```{r H4, echo=FALSE, cache=TRUE}
RT_size_resp <- inner_join(
  RT_size_present,
  RT_size_absent,
  by = "subj_id") %>%
  mutate(RT_size_resp = RT_size_present - RT_size_absent)
```

*Hypothesis 4 (SIZE RESPONSE INTERACTION ON REACTION TIME)*: We tested the null hypothesis that the effect of stimulus size on reaction time is similar in target-absent and target-present responses.

We performed a group-level t-test on a subject-level contrast in correct trials only: $(median({RT}_{P,small})-median({RT}_{P,big}))-(median({RT}_{A,small}) - median({RT}_{A,big}))$ where P and A stand for present and absent respectively.

This revealed a significant interaction between stimulus size and target presence (`r apa_print(RT_size_resp%>%pull(RT_size_resp)%>%t.test())$full_result`).

```{r H5, echo=FALSE, cache=TRUE}
accuracy_size <- size.df %>%
  dplyr::group_by(subj_id, size) %>%
  dplyr::summarise(
    hit_rate =(sum(correct & present)+0.5)/(sum(present)+1),
    fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
    dprime = qnorm(hit_rate)-qnorm(fa_rate),
    criterion = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))

dprime <- accuracy_size %>%
  dplyr::select(subj_id,size,dprime) %>%
  tidyr::pivot_wider(names_from=size, values_from=dprime) %>%
  dplyr::mutate(dprime_size=big-small)
```

*Hypothesis 5 (SENSITIVITY)*: We tested the null hypothesis that perceptual sensitivity (measured as $d\prime=z(H)-z(F)$) is equal as a function of the size of the stimulus. To allow the extraction of d’ for participants who committed no false-alarms or misses, we added 0.5 to miss, hit, false-alarm and correct rejection counts [@snodgrass2002].

We performed a paired t-test on the following subject-level contrast: $d\prime_{big} - d\prime_{small}$.

We found a significant effect of stimulus size on $d'$ (`r apa_print(dprime%>%pull(dprime_size)%>%t.test())$full_result`), with increased sensitivity for small stimuli.

```{r H6, echo=FALSE, cache=TRUE}
criterion <- accuracy_size %>%
  dplyr::select(subj_id,size,criterion) %>%
  tidyr::pivot_wider(names_from=size, values_from=criterion) %>%
  dplyr::mutate(criterion_size=big-small)
```

*Hypothesis 6 (CRITERION)*: We tested the null hypothesis that decision criterion (measured as $c=-0.5(z(H)+z(F))$) is unaffected by the size of the display. To allow the extraction of a decision criterion for participants who committed no false-alarms or misses, we added 0.5 to miss, hit, false-alarm and correct rejection counts [@snodgrass2002].

We performed a paired t-test on the following subject-level contrast: $c_{big} - c_{small}$.

We found a significantly more conservative criterion for big stimuli (`r apa_print(criterion%>%pull(criterion_size)%>%t.test())$full_result`).

```{r H7, echo=FALSE, cache=TRUE}
response_confidence <- size.df %>%
    dplyr::filter(correct == TRUE) %>%
    dplyr::group_by(subj_id, resp) %>%
    dplyr::summarise(confidence=mean(confidence)) %>%
    tidyr::pivot_wider(names_from = resp, values_from = confidence) %>%
    dplyr::mutate(response_confidence = `TRUE` - `FALSE`)
```

*Hypothesis 7 (PRESENCE/ABSENCE CONFIDENCE RATINGS)*: We tested the null hypothesis that confidence ratings are similar for target-present and target-absent responses, aiming to replicate the finding that subjective confidence levels are lower for decisions about absence than decisions about presence [@mazor2021].

We performed a paired t-test on the the following subject-level contrast from correct trials: $mean(confidence_{present}) - mean(confidence_{absent})$

This revealed a significant difference in confidence ratings for decisions about presence and absence (`r apa_print(response_confidence%>%pull(response_confidence)%>%t.test())$full_result`), with higher confidence ratings for present responses.

```{r H8, echo=FALSE, cache=TRUE}
conf_size_present <- size.df %>%
    dplyr::filter(present == 1 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(confidence = mean(confidence)) %>%
    tidyr::pivot_wider(names_from = size, values_from = confidence) %>%
    dplyr::mutate(conf_size_present = big-small)
```

*Hypothesis 8 (SIZE EFFECT ON CONFIDENCE IN PRESENCE)*: We tested the null hypothesis that target-present confidence ratings are similar when the stimulus is small or big.

We performed a paired t-test on the following individual-level contrast from correct trials: $mean(confidence_{P, big}) - mean(confidence)_{P, small}$ where P stands for present.

This revealed a significant difference in confidence ratings for big and small stimuli in target-present responses (`r apa_print(conf_size_present%>%pull(conf_size_present)%>%t.test())$full_result`), with higher confidence ratings for small stimuli.

```{r H9, echo=FALSE, cache=TRUE}
conf_size_absent <- size.df %>%
    dplyr::filter(present == 0 & correct == TRUE) %>%
    dplyr::group_by(subj_id, size) %>%
    dplyr::summarise(confidence = mean(confidence)) %>%
    tidyr::pivot_wider(names_from = size, values_from = confidence) %>%
    dplyr::mutate(conf_size_absent = big-small)
```

*Hypothesis 9 (SIZE EFFECT ON CONFIDENCE IN ABSENCE)*: We tested the null hypothesis that target-absent confidence ratings are similar when the stimulus is small or big.

We performed a paired t-test on the following individual-level contrast from correct trials: $mean(confidence_{A, big}) - mean(confidence)_{A, small}$ where A stands for absent.

This revealed a significant effect of stimulus size on confidence ratings in target-absent responses (`r apa_print(conf_size_absent%>%pull(conf_size_absent)%>%t.test())$full_result`), with higher confidence ratings for big stimuli.

```{r H10, echo=FALSE, cache=TRUE}
conf_size_resp <- inner_join(
  conf_size_present,
  conf_size_absent,
  by = "subj_id") %>%
  dplyr::mutate(conf_size_resp = conf_size_present - conf_size_absent)
```

*Hypothesis 10 (SIZE RESPONSE INTERACTION ON CONFIDENCE)*: We tested the null hypothesis that the effect of stimulus size on confidence ratings is similar in target-present and target-absent responses.

We performed a group-level t-test on a subject-level contrast in correct trials only: $(mean({confidence}_{P,small})-mean({confidence}_{P,big}))-(mean({confidence}_{A,small}) -mean({confidence}_{A,big}))$ where P and A stand for present and absent respectively.

This revealed a significant interaction effect between stimulus size and target presence on subjective confidence levels (`r apa_print(conf_size_resp%>%pull(conf_size_resp)%>%t.test())$full_result`).

\newpage

## References
