---
title: "Results section"
output: word_document
date: "2025-05-02"
bibliography: references.bib
---


```{r global_options, include=FALSE, cache=F}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, cache=T, sanitize = T)
```

```{r setup, include = FALSE, cache=F}

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse',
    'BayesFactor',# for Bayesian t test
    'ggrepel', # for geom_label_repel
    'lmerTest',
    'osfr'
  ), "2024-12-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)

```

```{r download_load_and_format_data, echo=FALSE, cache=TRUE}

source("../analysis/downloadFromOSF.R")
source("../analysis/loadAndPreprocessData.R")

```

```{r overall-descriptives, echo=FALSE, cache=TRUE}

size.overall_descriptives<- size.df %>%
  group_by(subj_id) %>%
  filter(test_part %in% c('test1','test2'))%>%
  summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))


size.descriptives_by_size<- size.df %>%
  group_by(subj_id,size) %>%
  filter(test_part %in% c('test1','test2'))%>%
  summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            pcorrect_yes = sum(correct & present)/(sum(resp)),
            pcorrect_no = sum(correct & !present)/(sum(!resp)),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))

size.d_by_size <- size.descriptives_by_size %>%
  dplyr::select (subj_id,size,d)%>%
  spread(size,d)%>%
  mutate(diff=small-big)

size.present_correct_RT_by_size <- size.df %>%
  group_by(subj_id,size) %>%
  filter(test_part %in% c('test1','test2') & present & correct) %>%
  summarise(RT=median(RT)) %>%
  spread(size,RT) %>%
  mutate(diff=big-small)

size.present_correct_confidence_by_size <- size.df %>%
  group_by(subj_id,size) %>%
  filter(test_part %in% c('test1','test2') & present & correct) %>%
  summarise(confidence=mean(confidence)) %>%
  spread(size,confidence) %>%
  mutate(diff=big-small)

size.absent_correct_RT_by_size <- size.df %>%
  group_by(subj_id,size) %>%
  filter(test_part %in% c('test1','test2') & !present & correct) %>%
  summarise(RT=median(RT)) %>%
  spread(size,RT) %>%
  mutate(diff=big-small)

size.absent_correct_confidence_by_size <- size.df %>%
  group_by(subj_id,size) %>%
  filter(test_part %in% c('test1','test2') & !present & correct) %>%
  summarise(confidence=mean(confidence)) %>%
  spread(size,confidence) %>%
  mutate(diff=big-small)

```

`r size.raw_df$subj_id%>%unique()%>%length()` English-speaking adult participants were recruited via Prolific.com to take part in Exp. 1 (pre-registration: [osf.io/r62nm](https://osf.io/r62nm)), in which they detected the presence or absence of a target letter (A or S, on different trials) in dynamic patches of visual noise. The target letter was present on 50% of the trials, and in the remaining 50% target-absent trials stimuli consisted of pure noise, generated by sampling, with replacement, pixels from the target image. The stimulus appeared on the screen, refreshing at 15 frames per second, until a response was recorded, at which point the participant rated their decision confidence on an analog scale. After applying our pre-registered exclusion criteria (above-chance accuracy, and no more than 25% of trials slower than 5 seconds or faster than 100 milliseconds), `r size.df$subj_id%>%unique()%>%length()` participants were included in the analysis. On average, participants responded correctly on `r size.overall_descriptives$accuracy%>%mean()%>%printnum()` (SD=`r size.overall_descriptives$accuracy%>%sd()%>%printnum()`) of trials, and had a slight but significant bias to report target absence (proportion of "target presence" responses: `r apa_print(size.overall_descriptives$resp_bias%>%t.test(mu=0.5))$full_result`).

Our critical manipulation was of stimulus size. Stimuli were 18-by-12 pixels in size, but on different trials individual pixels appeared large (5-by-5 screen pixels per stimulus pixel) or small (3-by-3 screen pixels per stimulus pixel). Letters were easier to detect when stimuli appeared smaller (d' for small stimuli: `r size.d_by_size$small%>%mean()%>%printnum()`; d' for big stimuli: `r size.d_by_size$big%>%mean()%>%printnum()`; difference: `r apa_print(size.d_by_size$diff%>%t.test)$statistic`), perhaps due to the spatial low frequency, pixelated nature of stimuli [@schyns1994]. Accordingly, participants were faster to detect letters when they appeared small (mean median reaction time in correct letter detections in seconds: small: `r (size.present_correct_RT_by_size$small%>%mean()/1000)%>%printnum()` seconds; big: `r (size.present_correct_RT_by_size$big%>%mean()/1000)%>%printnum()` seconds; `r apa_print(size.present_correct_RT_by_size$diff%>%t.test())$statistic`). They were also more confident in making decisions about presence when stimuli appeared small (mean confidence in correct letter detections on a 0-1 scale: small: `r size.present_correct_confidence_by_size$small%>%mean()%>%printnum()`; big: `r size.present_correct_confidence_by_size$big%>%mean()%>%printnum()`; `r apa_print(size.present_correct_confidence_by_size$diff%>%t.test())$statistic`), in line with confidence in presence tracking the objective visibility of stimuli. Finally, the three effects were correlated: participants whose perceptual sensitivity benefited more when stimuli appeared smaller were also relatively more confident (`r apa_print(cor.test(-size.d_by_size$diff, size.present_correct_confidence_by_size$diff))$full_result`) and faster (`r apa_print(cor.test(-size.d_by_size$diff, size.present_correct_RT_by_size$diff))$full_result`) in their correct letter detections in small compared to big displays.

```{r perceived-harder, echo=FALSE, cache=TRUE}
size.perceived_harder <- size.df %>%
  filter(trial_type=='survey-multi-choice' & grepl('which_was_harder',response)) %>%
  mutate(perceived_harder = ifelse(
    grepl('bigger',response),
    'big',
    ifelse(grepl('smaller',response),
           'small',
           'no-difference'))) %>%
  dplyr::select(subj_id,perceived_harder)

size.d_by_size_with_report <- size.d_by_size %>%
  merge(size.perceived_harder)

size.absent_correct_confidence_by_size_with_report <- size.absent_correct_confidence_by_size %>%
  merge(size.perceived_harder)
  
```

Critically, however, this negative effect of stimulus size on target visibility went against participants' reported beliefs. When asked at the end of the experiment -- that is, after having experienced that small stimuli are easier to detect -- whether stimulus size made any difference to task difficulty, only a small minority of `r size.perceived_harder%>%filter(perceived_harder=='big')%>%nrow()`/`r size.perceived_harder%>%nrow()` participants reported that bigger letters were harder to detect, `r size.perceived_harder%>%filter(perceived_harder=='no-difference')%>%nrow()` participants reported no difference, and the remaining `r size.perceived_harder%>%filter(perceived_harder=='small')%>%nrow()` reported that smaller stimuli were harder to detect. Furthermore, there was no association between reported beliefs and the objective effect of size on their perceptual sensitivity across participants (`r apa_print(summary(aov(diff~perceived_harder,data=size.d_by_size_with_report)))$statistic`). Exp. 2 below provides additional evidence that participants erroneously believed bigger stimuli were easier to detect, using prospective confidence ratings. Together, our design fully dissociated between objective and perceived difficulty: while detection was objectively easier when stimuli appeared small, participants believed the opposite was the case.

### Confidence in absence tracks perceived, not objective, visibility

```{r detection-asymmetries, echo=FALSE, cache=TRUE}

size.RT_by_response <- size.df %>%
  group_by(subj_id, present) %>%
  filter(test_part %in% c('test1','test2') & correct)%>%
  summarise(RT=median(RT))%>%
  spread(present,RT,sep='')%>%
  mutate(diff=present0-present1)

size.confidence_by_response <- size.df %>%
  group_by(subj_id, present) %>%
  filter(test_part %in% c('test1','test2') & correct)%>%
  summarise(confidence=mean(confidence))%>%
  spread(present,confidence,sep='')%>%
  mutate(diff=present1-present0)
```

With this dissociation, we returned to decisions about absence. As per our logic, these decisions are made not upon perceiving the absence of a stimulus, but based on a failure to perceive their presence, together with a belief that a target would have been perceived if it was present [@mazor2025]. In line with absence being inferred tentatively and based on the absence of evidence, we replicated detection asymmetries in reaction time (mean median reaction times in decisions about absence: `r (size.RT_by_response$present0%>%mean()/1000)%>%printnum()` seconds; decisions about presence: `r (size.RT_by_response$present1%>%mean()/1000)%>%printnum()` seconds; mean difference: `r (size.RT_by_response$diff%>%mean()/1000)%>%printnum()` seconds; `r apa_print(size.RT_by_response$diff%>%t.test())$statistic`) and confidence (mean confidence in decisions about absence: `r (size.confidence_by_response$present0%>%mean())%>%printnum()`; decisions about presence: `r (size.confidence_by_response$present1%>%mean())%>%printnum()`; mean difference: `r (size.confidence_by_response$diff%>%mean())%>%printnum()` on a 0-1 scale; `r apa_print(size.confidence_by_response$diff%>%t.test())$statistic`).

Importantly, if decisions about absence are made based on a belief about visibility, rather than visibility itself, confidence in those trials should not be higher when visibility was high (small stimuli), but when visibility was believed to be high in participants' internal model of their own perception (big stimuli; "It would have been hard to see the target in this small display, so I might have missed it"). Indeed, unlike confidence in presence, which was higher for smaller stimuli, confidence in absence was higher for bigger stimuli (mean confidence in correct rejections: small: `r size.absent_correct_confidence_by_size$small%>%mean()%>%printnum()`; big: `r size.absent_correct_confidence_by_size$big%>%mean()%>%printnum()`; `r apa_print(size.absent_correct_confidence_by_size$diff%>%t.test())$statistic`). Furthermore, unlike the effect of stimulus size on confidence in presence, here we observed no correlation with the effect of stimulus size on perceptual sensitivity (`r apa_print(cor.test(size.d_by_size$diff, size.absent_correct_confidence_by_size$diff))$full_result`). Participants' confidence in absence was determined by their inaccurate beliefs about visibility, rather than by visibility itself.

```{r create-figure-1, echo=FALSE, cache=TRUE}

detection_colors <- c("#E41a1c","#377eb8");

size.descriptives_by_size  %>%
  group_by(size) %>%
  summarise(se=se(d), d=mean(d)) %>%
  ggplot(aes(x=size,y=d,ymin=d-se,ymax=d+se))+
  geom_point(size=2)+
  geom_errorbar()+
  labs(x='',y="sensitivity (d')")+
  theme_classic()

ggsave('figures/Exp1_dprime.png',width=1.8,height=2.5,dpi=300)

size.df %>%
  filter(correct & test_part %in% c('test1','test2')) %>%
  group_by(subj_id,present,size) %>%
  summarise(confidence=mean(confidence))%>%
  group_by(present,size)%>%
  summarise(se=se(confidence),confidence=mean(confidence)) %>%
  mutate(present=factor(present, levels=c(0,1),labels=c('absent','present'))) %>%
  ggplot(aes(x=size,y=confidence,color=present,group=present, ymin=confidence-se,ymax=confidence+se))+
  geom_line()+
  geom_point(size=2)+
  geom_errorbar()+
  scale_color_manual(values=detection_colors)+
  labs(x='',y='mean confidence')+
  theme_classic()+
  theme(legend.position='none')

ggsave('figures/Exp1_confidence.png',width=2.2,height=2.5,dpi=300)

size.d_by_size %>%
  mutate(d_diff=-diff)%>% #so that everything is big-small
  dplyr::select(subj_id,d_diff)%>%
  merge(size.absent_correct_confidence_by_size %>%
          mutate(absent=diff)%>%
          dplyr::select(subj_id,absent)) %>%
  merge(size.present_correct_confidence_by_size %>%
          mutate(present=diff)%>%
          dplyr::select(subj_id,present)) %>%
  pivot_longer(cols=c(present, absent),
               names_to='present',
               values_to='conf_diff') %>%
  ggplot(aes(x=d_diff,y=conf_diff,color=present))+
  geom_point(alpha=0.3)+
  geom_smooth(method = "lm", se=FALSE)+
  facet_wrap(~present)+
  scale_color_manual(values=detection_colors)+
  labs(x= "Δd' (big−small)",y="Δconfidence (big−small)")+
  theme_classic()+
  theme(legend.position='none')
  
ggsave('figures/Exp1_correlations.png',width=4,height=2.5,dpi=300)


```

### An ideal observer model of perceptual detection dissociates between true and believed visibility

As shown above, our paradigm successfully dissociated between true and believed visibility. A recent model of perceptual detection allows to formalise this dissociation, describing perceptual detection as a series of momentary decisions made by an ideal observer in a partially observed environment [for a more technical description of the model, see @mazor2025]. On each trial, the agent is in one of two world states: a target letter is either objectively present, or absent. The agent does not have direct access to the world: it only has access to the noisy all-or-none activations of a sensor, which activates with probability $\theta_{absent}$ when a target is absent, and with a higher probability $\theta_{present}$ when a target is present. The agent observes a sequence of activations and inactivations at a rate of 20 per second, and, based on this information, decides whether to commit to a "target present" or a "target absent" decision, or, crucially, to accumulate more evidence. The trial ends once the agent commits to a decision, and the agent is internally rewarded for being correct.

Critically, the agent does not know the true values of $\theta_{absent}$ and $\theta_{present}$. Instead, it has subjective beliefs about these values $\bar{\theta}_{absent}$ and $\bar{\theta}_{present}$, which may be more or less accurate. This way, the model can describe a setting where perceptual information is of low quality ($\theta_{absent} \approx \theta_{present}$), but the agent believes it is of high quality ($\bar{\theta}_{absent}<\bar{\theta}_{present}$) or vice versa. Finally, in deciding how to act, the agent follows the optimal policy to maximise its long-term reward, based on its beliefs about visibility and subjective temporal discounting of value (which is assumed to take an exponential form). We derive the optimal policy by solving the Bellman equation using backward induction. As a consequence, this model does not have free parameters that correspond to decision biases, criterions or boundaries. Instead, model parameters correspond to properties of the sensor, the agents' beliefs about the sensor, and the agents’ value function and decision noise. The model is fit to decisions and decision times, but qualitative predictions about confidence can be made by taking the model-derived subjective probability of being correct at the time of making the decision.

Finally, decreasing the size of stimuli scales the probability of sensor activation by a factor of $\alpha$, corresponding to the true perceptual effect of size on visibility. The agent holds a subjective belief about this value, $\bar{\alpha}$, which they use to flexibly interpret evidence and set their subjective decision boundary as a function of stimulus size.

Ten model parameters were fitted to the data of individual subjects, with the parameters of interest being $\alpha$ and $\bar{\alpha}$, respectively corresponding to the true and believed effects of size on stimulus visibility. These were allowed to vary between 0.58 to 2.22. To account for noise in the decision-making process, action selection followed a softmax distribution. A previous study that used this model showed good parameter recovery [@mazor2025]. The model was fitted to the data of 


```{r load-fitted-parameters-and-simulated-data, echo=FALSE, cache=TRUE}
size.parameters <- read.table('../model/modelFitting/bestParameters/size/best_parameters_from_E_size_fit.csv', header=FALSE, sep=',') %>%
  mutate(alpha=V8**2,
         belalpha=V9**2,
         gamma = V5,
         diffalpha = belalpha-alpha)

size.sim.df <- read.csv('../model/modelFitting/simulateDataFromParameters/simulated_data/Simulated_data_E_size.csv') %>%
  mutate(present=factor(present,levels=c(1,-1),labels=c('present','absent')),
         size=ifelse(easy,5,3));

size.minimal.df <- read.csv('../model/data/E_size_2.csv') %>%
  mutate(present=factor(present,levels=c(1,-1),labels=c('present','absent')),
         size=ifelse(easy,5,3));
```

```{r simulated-and-real-confidence-effects, echo=FALSE, cache=TRUE}

size.confidence_effects <- size.minimal.df%>%
  dplyr::group_by(subj_id,present) %>%
  dplyr::summarise(diff = (mean(confidence[correct==1 & easy==1])-mean(confidence[correct==1 & easy==0]))/sd(confidence[correct==1]))%>%
  spread(key=present,value=diff)

size.sim_conf_effects <- size.sim.df %>%
  filter(correct==1)%>%
  group_by(subj_id,present) %>%
  dplyr::summarise(diff = (mean(confidence[correct==1 & easy==1])-mean(confidence[correct==1 & easy==0]))/sd(confidence[correct==1]))%>%
  spread(key=present,value=diff)

```

To examine the model’s success in capturing qualitative patterns found in behavioural data, we simulated data from the fitted model parameters. The simulated data accurately matched participants’ actual response time and error rates, predicting an advantage for smaller stimuli when a target is present, but not when a target is absent (see Fig. 3B, left and middle panels). In contrast, while the model correctly predicted the increase in confidence for for smaller stimuli in target-present trials, it failed to predict that the pattern should be opposite in target-absent trials (Fig. 3B). 

We next examined the fitted parameters. Across participants, $\alpha$ was `r size.parameters$alpha%>%mean()%>%printnum()` (SD=`r size.parameters$alpha%>%sd()%>%printnum()`) and significantly lower than 1 (a t-test on log($\alpha$) against 0: `r apa_print(size.parameters$alpha%>%log()%>%t.test())$statistic`), meaning that, according to the model, big stimuli were objectively less visible . Surprisingly, $\bar\alpha$ was also slightly lower than 1 (mean=`r size.parameters$belalpha%>%mean()%>%printnum()`, SD=`r size.parameters$belalpha%>%sd()%>%printnum()`; a t-test on log($\bar\alpha$) against 0: `r apa_print(size.parameters$belalpha%>%log()%>%t.test())$statistic`), meaning that, according to the model, big stimuli were also believed to be less visible. This goes against participants’ self-reported difficulty ratings and confidence in absence in Exp.1.  Critically however, the model correctly produced a significant difference between $\alpha$ and $\bar\alpha$  across individuals (`r apa_print(t.test(size.parameters%>%mutate(diff=log(alpha)-log(belalpha))%>%pull(diff)))$statistic`), thereby capturing the fact that participants systematically misrepresented the effect of stimulus size on stimulus visibility.

```{r correlation-between-parameters-and-data, echo=FALSE, cache=TRUE}
corr_data <- bind_cols(size.parameters, size.confidence_effects) %>%
  select(alpha,belalpha,present, absent) 
```

Some support for our interpretation of these models parameters as reflecting true and believed visibility comes from their correlations with confidence effects. $\alpha$, which reflects the effect of size on stimulus visibility, was strongly correlated with the effect of size on confidence in target-presence (`r apa_print(cor.test(corr_data$alpha, corr_data$present,method='spearman'))$full_result`), but not at all in target-absence (`r apa_print(cor.test(corr_data$alpha, corr_data$absent,method='spearman'))$full_result`; Fig. 4, first row). Conversely, $\bar\alpha$, which reflects participants’ beliefs about the effect of size on stimulus visibility, was not at all correlated with the effect of size on confidence in presence (`r apa_print(cor.test(corr_data$belalpha, corr_data$present,method='spearman'))$full_result`), but it correlated with the effect of size on confidence in absence (`r apa_print(cor.test(corr_data$belalpha, corr_data$absent, method='spearman'))$full_result`). Notably, these are correlations between model parameters and out-of-sample data which was not available to the model during the parameter fitting process (the model was fitted to decisions and decision times, not confidence ratings). This double dissociation supports our psychological interpretation of these two model parameters as reflecting true and believed effects of size on visibility, as well as our broader theoretical claim that confidence in presence reflects true visibility and confidence in absence reflects believed, or counterfactual, visibility.

### Time-resolved analysis reveals meta-perceptaul learning within the experiment

```{r time-resolved, echo=FALSE, cache=TRUE}

detection_colors <- c("#E41a1c","#377eb8");

size.conf_effect_by_block <- size.df %>%
  filter(test_part %in% c('test1','test2')) %>%
  group_by(subj_id) %>%
  mutate(trial_index = 1:n(),
         block = floor((trial_index-1)/16)+1)%>%
  filter(correct==1)%>%
  group_by(subj_id,block, present) %>%
  summarise(diff=mean(confidence[size=='big'])-mean(confidence[size=='small']))

size.dprime_by_block <- size.df %>%
  filter(test_part %in% c('test1','test2')) %>%
  group_by(subj_id) %>%
  mutate(trial_index = 1:n(),
         block = floor((trial_index-1)/16)+1)%>%
  group_by(subj_id,block, size) %>%
  summarise(hit_rate = (sum(correct==1 & present==1)+0.5)/(sum(present==1)+1),
            fa_rate = (sum(correct==0 & present==0)+0.5)/(sum(present==0)+1),
            d = qnorm(hit_rate)-qnorm(fa_rate))%>%
  group_by(subj_id,block)%>%
  summarise(diff=d[size=='big']-d[size=='small'])

size.conf_effect_by_block %>%
  group_by(block,present)%>%
  summarise(se=se(diff),diff=mean(diff, na.rm=TRUE)) %>%
  mutate(present=factor(present, levels=c(0,1),labels=c('absent','present'))) %>%
  ggplot(aes(x=block,y=diff,color=present,group=present, ymin=diff-se,ymax=diff+se))+
  geom_abline(slope=0,intercept=0,linetype=2)+
  geom_line()+
  geom_point(size=2)+
  geom_errorbar()+
  scale_color_manual(values=detection_colors)+
  labs(x='block number',y='Δconfidence (big−small)')+
  theme_classic()+
  theme(legend.position='none')

ggsave('figures/Exp1_conf_by_block.png',width=3.2,height=2.5,dpi=300)

size.block_effect_present <- lmer(diff~block+(1|subj_id),data=size.conf_effect_by_block%>%filter(present==1))
size.block_effect_absent <- lmer(diff~block+(1|subj_id),data=size.conf_effect_by_block%>%filter(present==0))
size.block_effect_dprime <- lmer(diff~block+(1|subj_id),data=size.dprime_by_block)
size.block_effect_interaction <- lmer(diff~block*present+(1|subj_id),data=size.conf_effect_by_block%>%mutate(diff=ifelse(present==1,diff,-diff)))

size.learning_slopes <-size.conf_effect_by_block %>% 
  group_by(present,subj_id) %>% 
  summarise(slope=coef(lm(diff~block))[2]) %>%
  spread(present,slope,sep='')

size.dprime_slopes <- size.dprime_by_block %>% 
  group_by(subj_id) %>% 
  summarise(dprime=coef(lm(diff~block))[2]) 

size.learning_slopes <- size.learning_slopes %>%
  merge(size.dprime_slopes)
```

If participants started the experiment expecting bigger letters to be easier to detect, by the end of it they had experienced 15 minutes of feedback to the contrary, such that small letters were systematically easier to detect. In our next analysis, we asked whether experience with the task resulted in learning about the true effect of size on perceptual sensitivity, and whether this was evident in confidence ratings. We therefore computed the difference in confidence in correct responses as a function of stimulus size, block number (each block consisted of 16 trials) and target presence (see Fig. 5). 

A clear pattern emerged: the effect of size on confidence systematically decreased as a function of block number, both when a target was absent (`r apa_print(anova(size.block_effect_absent))$full_result$block` for the fixed effect in the model $\Delta confidence\sim block+(1|subject)$), and, surprisingly, when a target was present (although this decrease was not as steep; `r apa_print(anova(size.block_effect_present))$full_result$block`). While our model can account for the target-absent effect as reflecting meta-perceptual learning, the target-present effect is harder to explain, especially given that the true effect of stimulus size on visibility, measured with detection sensitivity, remained similar between blocks `r apa_print(size.dprime_slopes$dprime%>%t.test())$statistic`. Tellingly, the two effects were statistically independent: participants who showed the steepest drop in effect size across blocks in target-absent trials did not necessarily show a similar drop in target-present trials (correlation between slopes from the model $\Delta confidence \sim block$ fitted to target-present and target-absent trials of individual subjects: `r apa_print(cor.test(size.learning_slopes$present0,size.learning_slopes$present1))$full_result`), suggestive of a different mechanism behind the two effects. We return to this point in the Discussion.

### Prospective confidence ratings follow beliefs about visibility 

We reasoned that if confidence in absence reflects participants’ beliefs about perception, rather than anything about the perceptual input itself, it should align with participants’ confidence prior to seeing the stimulus at all. In Exp. 2, we tested this hypothesis. 
`r prospective.raw$subj_id%>%unique()%>%length()` English-speaking adult participants were recruited via Prolific.com to participate in Exp. 2 (pre-registration: [osf.io/z652f](https://osf.io/z652f)). After applying our pre-registered exclusion criteria (above-chance accuracy, and no more than 25% of trials slower than 5 seconds or faster than 100 milliseconds), `r prospective.df$subj_id%>%unique()%>%length()` participants were included in the analysis.

Similar to Exp.1, the critical manipulation was one of stimulus size, with stimuli appearing either large or small on different trials. However, this time, in the first two blocks of the experiment we asked participants to rate their confidence before deciding whether the target is present or absent. To this end, participants were presented with an empty black frame indicating the size of the stimulus they were about to observe and asked to rate how confident they are that they will be correct in the upcoming decision. Only afterwards were they presented with the noisy stimulus and asked to detect the presence or absence of the target (Fig. 6A, first row). Blocks 3 and 4 followed a similar trial structure to that of Exp. 1 (Fig. 6A, second row). Since our time-resolved analysis from Exp. 1 revealed rapid meta-perceptual learning within 64 experimental trials, and since we were most interested in participants’ miscalibrated beliefs about their perception, Exp. 1 included only 32 trials: 16 with prospective and 16 with post-decisional confidence ratings. 

```{r prospective-analysis, echo=FALSE, cache=TRUE}

prospective.overall_descriptives<- prospective.df %>%
  group_by(subj_id) %>%
  summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))


prospective.descriptives_by_size<- prospective.df %>%
  group_by(subj_id,size) %>%
  summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            pcorrect_yes = sum(correct & present)/(sum(resp)),
            pcorrect_no = sum(correct & !present)/(sum(!resp)),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))

prospective.d_by_size <- prospective.descriptives_by_size %>%
  dplyr::select (subj_id,size,d)%>%
  spread(size,d)%>%
  mutate(diff=small-big)

prospective.prospective_confidence_by_size <- prospective.df %>%
  filter(confidence_time=='before' & correct==1) %>%
  group_by(subj_id,size) %>%
  summarise(confidence=mean(confidence)) %>%
  spread(size,confidence) %>%
  mutate(diff=big-small)

prospective.present_confidence_by_size <- prospective.df %>%
  filter(confidence_time=='after' & correct==1 & present==1) %>%
  group_by(subj_id,size) %>%
  summarise(confidence=mean(confidence)) %>%
  spread(size,confidence) %>%
  mutate(diff=big-small)

prospective.absent_confidence_by_size <- prospective.df %>%
  filter(confidence_time=='after' & correct==1 & present==0) %>%
  group_by(subj_id,size) %>%
  summarise(confidence=mean(confidence)) %>%
  spread(size,confidence) %>%
  mutate(diff=big-small)

### RT

prospective.RT_present <- prospective.df %>% 
  dplyr::filter(RT > 100 & RT < 7000 & present == 1 & correct == TRUE) %>%
  dplyr::group_by(subj_id, size) %>%
  dplyr::summarise(RT = median(RT))%>%
  tidyr::pivot_wider(names_from = size, values_from = RT) %>%
  dplyr::mutate(diff = `big` - `small`)

# H3
prospective.RT_absent <- prospective.df %>%
  dplyr::filter(RT > 100 & RT < 7000 & present == 0 & correct == TRUE) %>%
  dplyr::group_by(subj_id, size) %>%
  dplyr::summarise(RT = median(RT))%>%
  tidyr::pivot_wider(names_from = size, values_from = RT) %>%
  dplyr::mutate(diff = `big` - `small`)

```


Here, too, letters were easier to detect when the stimuli appeared smaller (d’ for small stimuli: `r prospective.d_by_size$small%>%mean()%>%printnum()`, d’ for big stimuli: `r prospective.d_by_size$big%>%mean()%>%printnum()`; Fig. 6B). Critically, prospective confidence ratings showed the opposite pattern, with participants being significantly more confident that they will be correct in their upcoming decision when they expected the stimulus to be big, compared to small  (mean confidence in upcoming letter detections on a 0-1 scale: small: `r prospective.prospective_confidence_by_size$small%>%mean()%>%printnum()`; big: `r prospective.prospective_confidence_by_size$big%>%mean()%>%printnum()`; `r apa_print(prospective.prospective_confidence_by_size$diff%>%t.test())$statistic`; Fig. 6C). This is consistent with a prior meta-perceptual belief that bigger stimuli are more visible. 
Confidence in absence showed a similar pattern to prospective confidence (in full dissociation from objective accuracy), and was higher when stimuli were expected to appear big (mean confidence in correct rejections: small: `r prospective.absent_confidence_by_size$small%>%mean()%>%printnum()`, big: `r prospective.absent_confidence_by_size$big%>%mean(na.rm=T)%>%printnum()`; `r apa_print(prospective.absent_confidence_by_size$diff%>%t.test())$statistic`). Intriguingly, confidence in presence was unaffected by the size manipulation (mean confidence in correct letter detections: small: `r prospective.present_confidence_by_size$small%>%mean()%>%printnum()`, big: `r prospective.present_confidence_by_size$big%>%mean(na.rm=T)%>%printnum()`; `r apa_print(prospective.present_confidence_by_size$diff%>%t.test())$statistic`), and so were target-present reaction times (mean median reaction time in correct letter detections in seconds: small: `r (prospective.RT_present$small%>%mean()/1000)%>%printnum()`, big: `r (prospective.RT_present$big%>%mean()/1000)%>%printnum()`, `r apa_print(prospective.RT_present$diff%>%t.test())$statistic`). Finally, across subjects, the effect of stimulus size on post-decisional confidence ratings was uncorrelated with the effect of stimulus size on perceptual sensitivity or with the effect of stimulus size on prospective confidence, both in target-present (all p’s>.30) and in target-absent trials (all p’s>.62). 



::: {#refs custom-style="Bibliography"}
:::
